---
title: "GLM MC"
author: "Brooke Walters"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidymodels)
library(tidyverse)
```


```{r message=FALSE, warning=FALSE}
library(readr)
inventory1 <- read_csv("/home/knightsa/STA_631/Activities/activity08-mini-competition/competition-files/inventory.csv")
```

make new variables
categorize the data into seasons
```{r}

# Add a new column to indicate which chunk of 9 weeks each week belongs to
inventory1 <- inventory1 %>%
  mutate(sales_period = cut(week, breaks = seq(0, 54, 9), labels = FALSE, include.lowest = TRUE))

```


create a totals sales column for comparison (to be omitted later)
```{r}
# Calculate total sales for each item in each chunk
salesp<- inventory1 %>%
  group_by(item_no, sales_period) %>%
  summarise(total_sold = sum(sold, na.rm = TRUE), .groups = "drop")

```


# create an average sales column for comparison (to be omitted later)
```{r}
# Calculate the yearly average for each item
average_sales <- salesp %>%
  group_by(item_no) %>%
  summarise(average_yearly_sales = mean(total_sold))

# Join this back to the original data frame
salesp <- left_join(salesp, average_sales, by = "item_no")
```


create a total sales column (to be omitted later)
```{r}
# Calculate the yearly total for each item
total_sales <- salesp %>%
  group_by(item_no) %>%
  summarise(yearly_sales = sum(total_sold))

# Join this back to the original data frame
salesp <- left_join(salesp, total_sales, by = "item_no")
```


compare sales to the average (to be omitted later)
```{r}

# Define your threshold as 20% of the average
threshold <- 0.50

salesp <- salesp %>%
  mutate(
    season_to_avg = case_when(
      total_sold > (1 + threshold) * average_yearly_sales ~ "Significantly higher",
      total_sold < (1 - threshold) * average_yearly_sales ~ "Significantly lower",
      TRUE ~ "Average"
    )
  )

```


determines if the sales product is sold evenly thoughout the year
```{r}
# Creating "year_round_product" column
salesp <- salesp %>%
  group_by(item_no) %>%
  mutate(year_round_product = ifelse(sum(season_to_avg == "Average") >= 3, 1, 0)) %>%
  ungroup()
```


determines if the product is sold mostly during one point in the year 51% of the sales are sold in the one sales period
```{r}
# Create the 'main_season' column
salesp <- salesp %>%
  group_by(item_no) %>%
  mutate(in_seasonal_dominance = ifelse(total_sold / sum(total_sold) > 0.51, 1, 0)) %>%
  ungroup()

```



determines if the product is sold 40% in one sales period
```{r}
# Create the 'secondary_season' column
salesp <- salesp %>%
  group_by(item_no) %>%
  mutate(in_secondary_seasonal_dominance = ifelse(year_round_product == 0 & in_seasonal_dominance == 0 & total_sold / yearly_sales > 0.39, 1, 0)) %>%
  ungroup()

```


determins if the product makes up the top 10% of sales (best sellers)
```{r}
# Calculate the top 10% threshold
top_10_threshold <- quantile(salesp$yearly_sales, 0.9)

# Create the 'hv_product' column
salesp <- salesp %>%
  mutate(hv_product = ifelse(yearly_sales >= top_10_threshold, 1, 0))


```

```{r}
salesp %>% 
  filter(hv_product == 1) %>% 
  count(item_no)
```


```{r}
glimpse(salesp)
```

```{r}
salesp %>%
  ggplot(aes(x=sales_period, y=total_sold, color=factor(hv_product))) +
  geom_jitter(alpha = 0.4, size = 2)
```


```{r}
salesp1 <- salesp %>% 
  select(-average_yearly_sales, -yearly_sales, -season_to_avg)
```


```{r}
write.csv(salesp1, "wrangled_inventory.csv")
```


```{r}
head(salesp1)
```
```{r}
recipe <- recipe(total_sold ~ item_no + sales_period + year_round_product + in_seasonal_dominance + in_secondary_seasonal_dominance + hv_product, data = salesp1)

```

```{r}
model_spec <- linear_reg() %>% 
  set_engine("glm") %>% 
  set_mode("regression")
```


```{r}
workflow <- workflow() %>% 
  add_model(model_spec) %>% 
  add_recipe(recipe)
```

```{r}
model <- fit(workflow, data = salesp1)

model
```

```{r}
summary(model)
```
```{r}
model_fit <- extract_fit_parsnip(model)
```

```{r}
summary(model_fit)
```

```{r}
# Assuming your model is named 'model' and your data is stored in 'salesp1'

# Generate predicted values
predicted_values <- predict(model, new_data = salesp1)

# Combine the predicted values with the actual values
comparison <- data.frame(Actual = salesp1$total_sold, Predicted = predicted_values)

# Print the comparison
print(comparison)

```
```{r}
comparison %>% 
  filter(Actual == 0)
```

```{r}
prediction_table %>% 
  filter(total_sold == 0)
```


```{r}
prediction_table <- merge(x = salesp1, y = comparison, by.x = "total_sold" , by.y = "Actual")
```

```{r}
# reordering the columns so item sold is next to predicted

prediction_table <- prediction_table[,c(2:7, 1, 8)]

```

```{r}
#A dataset with the items our model actually predicts (No negative predictions)
 items_with_pos_prediction <- prediction_table %>% 
  mutate( diff = total_sold - .pred) %>% 
  filter(.pred > 0)
```

```{r}

items_with_pos_prediction <- items_with_pos_prediction %>%
  mutate(
    predict_status = case_when(
      diff > 0 ~ "Under Predict",
      diff < 0 ~ "Over Predict",
      diff == 0 ~ "Predicted"
    )
  )

```

```{r}
#eliminate diff column to avoid confusion

items_with_pos_prediction <- items_with_pos_prediction[,c(1:8,10)] 

```

